Simulations of Ballot Polling Risk-Limiting Audits

### 2022-10-11 providence/sections/workload.tex
acknowledgements, etc

Consequently, we now consider a model with a constant per-ballot cost $c_b$ and a constant per-round cost $c_r$.
So for an audit $\mathcal{A}$ with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we estimate that the cost $C$ of the audit is
\begin{equation}
C(\mathcal{A}) = E_b c_b + E_r c_r
\label{eq:round_cost}
\end{equation}

For simplicity, (and without loss of generality), we assume the per ballot cost is one, $c_b=1$. Then the per round cost $c_r$ tells us the cost of a round as a number of ballots. We begin with $c_r=1000$ as a conservative example. 
That is, we set the overhead of a round equal to the workload of sampling $1000$ ballots. Based on available data\cite{RI-report}, the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $c_r=1000$ is equivalent to roughly $20$ person-hours of workload. This corresponds to about $15$ minutes being spent per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate. 
As shown in Figure~\ref{fig:with_round_cost}, lower average costs are achieved by selecting higher stopping probability; \Providence achieves the lowest minimum average cost is achieved at roughly $0.7$.

\begin{figure}
\includegraphics[width=.5\textwidth]{with_round_cost.png}
\caption{For cost parameters $c_b=1$ and $c_r=1000$, this plot shows the expected cost for various round schedule parameters $p$. Expected cost is found using Equation~\ref{eq:cost} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
\label{fig:with_round_cost}
\end{figure}

Importantly, this gives us a way to estimate the expected cost, as well as which round schedule value $p$ achieves it, for arbitrary round cost. For each round cost $c_r$, we produce a dataset analagous to that of Figure~\ref{fig:with_round_cost} and then find the minimum average cost achieved for each of the audits and its corresponding stopping probability $p$. 
Figure~\ref{fig:optimal_costs} shows the optimal achievable workload for a wide range of per round costs. For very low round costs, the workload function approaches just the total number of ballots, and so, as seen in Figure~\ref{fig:abv_bals}, the three approaches differ by less. On the other hand, for extremely larger values of round cost, the average number of ballots has little impact on the workload function, and so the three audits again have similar values. For more reasonable values of the round cost $c_r$, SO \BRAVO and EoR \BRAVO achieve minimum cost roughly $1.1$ and $1.3$ times greater than that of \Providence.
Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve these minimal workloads. As expected, a overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audit $p$ increases as a function of $c_r$. Notice that \Providence is generally above and to the left of SO \BRAVO, and SO \BRAVO is generally above and to the left of EoR \BRAVO. This relationship reflects the fact that for the same round cost, \Providence can get away with a larger stopping probability because it requires fewer ballots.

\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_costs.png}
\caption{For varying round cost $c_r$, the optimal average cost achievable by each audit, as a fraction of the \Providence values. (We show the value for varying $c_r$ since the value of $c_r$ may differ significantly from case to case.) Note that values lower than $10^3$ are probably unrealistic in US statewide contests; in Virginia, $c_r=10^3$ corresponds to only about $15$ minutes of persontime per county as the overhead per round.}
\label{fig:optimal_costs}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_ps.png}
\caption{The optimal (cost-minimizing) stopping probability $p$ for varying cost model parameters $c_r$. With $c_b=1$, the varying value of $c_r$ can equivalently be thought of as the ratio of the cost of a round to the cost of a ballot. (Note that the steps in this function are a consequence of our subsampling the workload function. That is, the workload-minimizing value of $p$ for each $c_r$ is only allowed to take on values at increments of $0.5$.)}
\label{fig:optimal_ps}
\end{figure}

\noindent\textbf{Precinct overhead.} For a more complete model, we can also introduce container-level workload. The time to sample a ballot from an entirely new box is typically greater than to sample a ballot from an already-open box. Based on a Rhode Island pilot RLA report\cite{RI-report}, this may mean that a ballot from a new container requires roughly twice the time as a ballot from an already-opened container. Typically available election results give per-precinct granularity of vote tallies, rather than individual container information. In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload with an additional per-precinct constant cost, $c_p$. In this model, the workload estimate incurs an additional cost of $c_p$ every time a precinct is sampled from for the first time in a round. That is, let $E_{pi}$ be the expected number of distinct precincts sampled from in round $i$, and let $E_p=\sum_i E_{pi}$. Then the new model is
\begin{equation}
C(E_b, E_r, E_p) = E_b c_b + E_r c_r + E_p c_p
\label{eq:round_and_precinct_cost}
\end{equation}

We can again explore the minimum achievable workloads under this model, as shown in Figure~\ref{fig:optimal_cost_precinct_cost_ratio}.

\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_cost_precinct_cost_ratio.png}
\caption{The optimal average cost of the audits we simulated using workload function \ref{eq:round_and_precinct_cost} for varying $c_p$, given as a fraction of the value for \Providence}
\label{fig:optimal_cost_precinct_cost_ratio}
\end{figure}

\subsection{Real time}
Given tight certification deadlines\footnote{Virginia recently passed legislation requiring pre-certification RLAs TODO check exactly.}, the total real time to conduct the RLA is also an important factor to consider when planning audits.
Because each county can sample ballots for the same round concurrently, the total real time for a round depends only on the slowest county. 
In Virginia, Fairfax County typically has the most votes cast by a significant difference; in the contest we consider, Fairfax County had ~551 thousand votes cast, more than double the ~203 thousand of second-highest Virginia Beach City.
Consequently, we model the expected total real time $T$ of an audit using just the largest county, and we define analagous variables for the expected values in just the largest county.
For the largest county, let the expected total ballots sampled be $\bar E_b$, the expected number of rounds $\bar E_r$, and the expected number of distinct precinct samples summed over all rounds be $\bar E_c$.
Similarly, we use real time per-ballot, per-round, and per-precinct cost variables, $t_b$, $t_r$, and $t_p$. So the real time of the audit is estimated by
\begin{equation}
T(\bar E_b, \bar E_r, \bar E_p ) = \bar E_b t_b + \bar E_r t_r + \bar E_p t_p
\label{eq:real_time}
\end{equation}

@@

If election officials wish to enforce a \emph{misleading limit} for all the rounds, our simulation analysis could help. On the other hand, for a given round, it is straightforward to compute analytically the probability that a loser has more votes than the winner in the sample. Table~\ref{tab:misleading} shows for various margins the minimum first round size $n$ that guarantees a probability of a \emph{misleading sample} at most $M\in\{0.1,0.01,0.001\}$. For all values of $M$ and all margins, \Providence achieves a higher probability of stopping than either EoR \BRAVO or SO \BRAVO. 
    As seen in the Table~\ref{tab:misleading}, for $M=0.01$, the the minimum round sizes require at least roughly a $0.8$ probability of stopping in the first round. Even if the most efficient audit schedule (by either workload or real time measures) would use a lower stopping probability $p$, the election officials may use this constraint on the probability of a \emph{misleading sample} as the deciding factor in selecting a first round size.

>>>>

Consequently, we now consider a model with a constant per-ballot workload $w_b$ and a constant per-round workload $w_r$.
So for an audit with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we estimate that the workload $C$ of the audit is
\begin{equation}
W(E_b,E_r) = E_b w_b + E_r w_r
\label{eq:round_workload}
\end{equation}

For simplicity, (and without loss of generality), we assume the per ballot workload is one, $w_b=1$. Then the per round workload $w_r$ tells us the workload of a round as a number of ballots. We begin with $w_r=1000$ as a conservative example. 
That is, we set the overhead of a round equal to the workload of sampling $1000$ ballots. Based on available data\cite{RI-report}, the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $w_r=1000$ is equivalent to roughly $20$ person-hours of workload. This corresponds to about $15$ minutes being spent per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate. 
As shown in Figure~\ref{fig:with_round_workload}, lower average workloads are achieved by selecting higher stopping probability; \Providence achieves the lowest minimum average workload is achieved at roughly $0.7$.

\begin{figure}
\includegraphics[width=.5\textwidth]{with_round_workload.png}
\caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various round schedule parameters $p$. Expected workload is found using Equation~\ref{eq:workload} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
\label{fig:with_round_workload}
\end{figure}

Importantly, this gives us a way to estimate the expected workload, as well as which round schedule value $p$ achieves it, for arbitrary round workload. For each round workload $w_r$, we produce a dataset analagous to that of Figure~\ref{fig:with_round_workload} and then find the minimum average workload achieved for each of the audits and its corresponding stopping probability $p$. 
Figure~\ref{fig:optimal_workloads} shows the optimal achievable workload for a wide range of per round workloads. For very low round workloads, the workload function approaches just the total number of ballots, and so, as seen in Figure~\ref{fig:abv_bals}, the three approaches differ by less. On the other hand, for extremely larger values of round workload, the average number of ballots has little impact on the workload function, and so the three audits again have similar values. For more reasonable values of the round workload $w_r$, SO \BRAVO and EoR \BRAVO achieve minimum workload roughly $1.1$ and $1.3$ times greater than that of \Providence.
Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve these minimal workloads. As expected, a overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audit $p$ increases as a function of $w_r$. Notice that \Providence is generally above and to the left of SO \BRAVO, and SO \BRAVO is generally above and to the left of EoR \BRAVO. This relationship reflects the fact that for the same round workload, \Providence can get away with a larger stopping probability because it requires fewer ballots.

\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_workloads.png}
\caption{For varying round workload $w_r$, the optimal average workload achievable by each audit, as a fraction of the \Providence values. (We show the value for varying $w_r$ since the value of $w_r$ may differ significantly from case to case.) Note that values lower than $10^3$ are probably unrealistic in US statewide contests; in Virginia, $w_r=10^3$ corresponds to only about $15$ minutes of persontime per county as the overhead per round.}
\label{fig:optimal_workloads}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_ps.png}
\caption{The optimal (workload-minimizing) stopping probability $p$ for varying workload model parameters $w_r$. With $w_b=1$, the varying value of $w_r$ can equivalently be thought of as the ratio of the workload of a round to the workload of a ballot. (Note that the steps in this function are a consequence of our subsampling the workload function. That is, the workload-minimizing value of $p$ for each $w_r$ is only allowed to take on values at increments of $0.5$.)}
\label{fig:optimal_ps}
\end{figure}

\noindent\textbf{Precinct overhead.} For a more complete model, we can also introduce container-level workload. The time to sample a ballot from an entirely new box is typically greater than to sample a ballot from an already-open box. Based on a Rhode Island pilot RLA report\cite{RI-report}, this may mean that a ballot from a new container requires roughly twice the time as a ballot from an already-opened container. Typically available election results give per-precinct granularity of vote tallies, rather than individual container information. In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload with an additional per-precinct constant workload, $w_p$. In this model, the workload estimate incurs an additional workload of $w_p$ every time a precinct is sampled from for the first time in a round. That is, let $E_{pi}$ be the expected number of distinct precincts sampled from in round $i$, and let $E_p=\sum_i E_{pi}$. Then the new model is
\begin{equation}
W(E_b, E_r, E_p) = E_b w_b + E_r w_r + E_p w_p
\label{eq:round_and_precinct_workload}
\end{equation}

We can again explore the minimum achievable workloads under this model, as shown in Figure~\ref{fig:optimal_workload_precinct_workload_ratio}.

\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_workload_precinct_workload_ratio.png}
\caption{The optimal average workload of the audits we simulated using workload function \ref{eq:round_and_precinct_workload} for varying $w_p$, given as a fraction of the value for \Providence}
\label{fig:optimal_workload_precinct_workload_ratio}
\end{figure}

\subsection{Real time}
Given tight certification deadlines\footnote{Virginia recently passed legislation requiring pre-certification RLAs TODO check exactly.}, the total real time to conduct the RLA is also an important factor to consider when planning audits.
Because each county can sample ballots for the same round concurrently, the total real time for a round depends only on the slowest county. 
In Virginia, Fairfax County typically has the most votes cast by a significant difference; in the contest we consider, Fairfax County had ~551 thousand votes cast, more than double the ~203 thousand of second-highest Virginia Beach City.
Consequently, we model the expected total real time $T$ of an audit using just the largest county, and we define analagous variables for the expected values in just the largest county.
For the largest county, let the expected total ballots sampled be $\bar E_b$, the expected number of rounds $\bar E_r$, and the expected number of distinct precinct samples summed over all rounds be $\bar E_c$.
Similarly, we use real time per-ballot, per-round, and per-precinct workload variables, $t_b$, $t_r$, and $t_p$. So the real time of the audit is estimated by
\begin{equation}
T(\bar E_b, \bar E_r, \bar E_p ) = \bar E_b t_b + \bar E_r t_r + \bar E_p t_p
\label{eq:real_time}
\end{equation}

@@

If election officials wish to enforce a \emph{misleading limit} for all the rounds, our simulation analysis could help. On the other hand, for a given round, it is straightforward to compute analytically the probability that a loser has more votes than the winner in the sample. Table~\ref{tab:misleading} shows for various margins the minimum first round size $n$ that guarantees a probability of a \emph{misleading sample} at most $M\in\{0.1,0.01,0.001\}$. For all values of $M$ and all margins, \Providence achieves a higher probability of stopping than either EoR \BRAVO or SO \BRAVO. 
    As seen in the Table~\ref{tab:misleading}, to enforce $M=0.01$ requires minimum round sizes with at least roughly a $0.8$ probability of stopping in the first round. Even if the most efficient audit schedule (by either workload or real time measures) would use a lower stopping probability $p$ to choose the first round size, the election officials may opt to use this constraint on the probability of a \emph{misleading sample} as the deciding factor in planning their audits.
